# load and preprocessing the data

#### set my directory, url ect to download the data

```{r}
dir <- "D:/interest/coursera/Reproducible research/assignment1"
setwd(dir)
file.url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
```

#### download the data and unzip  
- just ignor the error in the chunk, for some reason url in the code is not compatible to Rmd

```{r,}
download.file(url = file.url, "assignment1.zip")
unzip("assignment1.zip")
```

#### read library for later analysis

```{r,  message=FALSE, warning=FALSE}
library(lubridate)
library(plyr)
library(ggplot2)
```

#### pre-process
- read the data
- transform the format
- delete the observation with NA

```{r}
data <- read.csv("activity.csv")
data$date <- ymd(as.character(data$date))
```

# What is mean total number of steps taken per day?

#### a histogram of the total number of steps taken each day

```{r}
data.per.day <- ddply(data, .(date), summarise, 
                      total.step = sum(steps, na.rm = T), 
                      mean.step = mean(steps, na.rm = T),
                      median.step = median(steps, na.rm = T))
hist(data.per.day$total.step)
```

#### mean and median total number of steps taken per day

```{r, results='hold'}
data.per.day[, c(1, 3, 4)]
```

# What is the average daily activity pattern?

####  time series plot on daily activity

```{r}
data.interval <- ddply(data, .(interval), summarise,
                       mean.step = mean(steps, na.rm = T))
with(data.interval, plot(interval, mean.step, type = "l"))
```

#### report the interval contain maxium number

```{r}
data.interval[which.max(data.interval$mean.step),]
```

# Imputing missing values

#### report the total number of missing values

```{r}
data.full <- read.csv("activity.csv")
number <- table(is.na(data.full$steps))["TRUE"]
print(as.numeric(number))
```

#### fill the data by daily mean of interval
#### create a new dataset with missing data fill in

```{r}
na.idx <- which(is.na(data.full$steps))
data.full$date <- ymd(data.full$date)
data.full[na.idx, ]$steps <- 
  data.interval[match(data.full[na.idx, ]$interval, data.interval$interval), "mean.step"]
```

#### compute daily steps

```{r}
data.full.daily <- ddply(data.full, .(date), summarise, 
                         mean.step = mean(steps), 
                         median.step = median(steps), 
                         total.step = sum(steps))
print(data.full.daily)
```

#### make a histograme and compare to previous one
#### the distribution looks more "normal"

```{r}
par(mfrow = c(1, 2))
hist(data.full.daily$total.step, col = "red", main = "after adding missing data")
hist(data.per.day$total.step, col = "blue", main = "original data")

```

# Are there differences in activity patterns between weekdays and weekends?

#### check the differnce between weekdays and weekends

```{r, fig.height = 8}
data$weekday <- weekdays(data$date)
data <- data[complete.cases(data),]
data$weekday.category <- NA
data$weekday.category <- ifelse(data$weekday %in% list("Sunday", "Saturday"), 
                            "weekend", "weekday")
data.weekday <- ddply(data, .(weekday.category, interval), summarise, 
                      mean.step = mean(steps))

```

#### make as plot for comparison

```{r}
ggplot(data.weekday, aes(interval, mean.step, color = weekday.category)) +
  geom_line() +
  facet_grid(weekday.category ~.)
```

